Research Experience

### 2010-06-24 doc/thesis/thesis.tex
more work on the exmaples, background knowledge, particle filter.

@@

\subsection{General Linear Model}
The most used form of FMRI analysis is Statistical Parametric
Mapping, but is able to account for several different levels or types
of stimulus (see \cite{Hofmann1997}). By using hierarchical models
the output signal timeseries is considered the weighted
sum of the various input timeseries. Essentially every experimental factor
is considered as another level inputs (ex. which patient is an input). 
The equation for a general linear model is then
\begin{equation}
Y(t) = X(t)\beta + \epsilon(t)
\end{equation}
where $Y(t)$ is the smoothed or detrended timeseries of measurements,
$X(t)$ is a row vector of input, $\beta$ is a column vector of weights,
and $\epsilon$ is the error. Thus for every time, the measurement is
assumed to be a weighted sum of the inputs plus some error. The calculation
of $\beta$ then performed using a maximum likelihood or gradient descent search 
to minimize the error.

@@

It is well known of course that a square wave stimulus does not result in a square wave
in the activation of brain regions. Thus, various methods are used to 
smooth $X(t)$'s related to actual input (as opposed to experimental confounds
such as patient number, age etc) through time, and bandlimit the input. 
The best technique is convolving the stimulus input with a hemodynamic 
response function, which mimicks the basic shape of BOLD activation, including a delay
due to rise time and fall time. The downside of this method, however is that 
the hemodynamic signal is static, meaning the same hemodynamic function is
used for every region of the brain, and $Y(t)$ must be a constant
scalar multiple of $X(t)$ to be appropiately identified.

@@

The GLM is extremely powerful at determining the linear dependence of
a set of regressors on the output. Unfortunately, there is significant evidence
that many of these dependencies are nonlinear, which means they may be
difficult or impossible to detect using linear techniques. Additionally, 
the lack of flexibility in delay variables certainly hinders SPM's
ability to locate voxels that do not conform to the canonical hemodynamic 
response function. The very fact that different hemodynamic response functions
work better for different regions/experiments is a testament to the need
for increased flexibility.

Linear models also ignore prior knowledge about the system. Combined FMRI
CBF or CBV imaging methods are available and could shed much light on
neural activation; however there is no way of combining that data into a
linear model. Although the most advanced form of the general linear model includes
a Hemodynamic Response Function, it is static across every region of the brain;
yet it is well known that capillary beds are not uniform and so blood perfusion 
cannot possibly be static across the brain. Its possible that activation is
every bit as strong in regions that aren't found by SPM, but that there are
nonlinearities that do not match the canonical HRF; resulting in signficant bias error. 
It is not uncommon
for data to be thrown out in FMRI studies because no significant activation
has been seen. This practice highlights the fact that strictly linear
approaches are insufficiently flexible to account for relatively common
variations of neural activation.

<Image with two different $\alpha$s>

<image comparing the results of 10\% changes in various signals>

@@

\begin{eqnarray}
X(t) = g(u(t), X(t-1))\\
Y(t) = h(X(t)
\end{eqnarray}

In order to estimate the posterior at $t$, a deterministic set of "sigma" points 
(often 2 per dimension, plus 1 at the mode of the multivariate distribution)
weighted according to a Gaussian estimate of $X(t-1)$ are passed through
the update equation. This set of points are then used to estimate the 
mean and covariance of $X(t)$. The benefit of this, is that it requires
no Jacobian and only a few extra calculations to get a decent esimate of
a posterior Gaussian. In the BOLD case, the set of equations we are modeling have no 
closed form solution, and finding the Jacobian is impossible without approximations. Although 
\cite{Riera2004} claims to use a Jacobian for their solution, they are
in fact using the Jacobian of $\dot{s}$, $\dot{q}$, $\dot{f}$, $\dot{v}$,
rather than $s$, $q$, $f$, and $v$. This is important because the Extended
Kalman filter depends on the jacobian to map a gaussian through the 
advancement of time. Thus the popular Extended Kalman Filter won't work
in this case, whereas the Unscented Kalman Filter still does.

@@

@@

\subsubsection{Resampling}
As a consequence 
of the wide prior distribution (required for a proper discretization of a continuous
distribution), there will be many particles with insignificant weights. While this does help
describe the tails of the distribution very well, it means that only a small portion of the
computation will be spent describing the most probable region. Ideally every particle would 
equally decrease the entropy of the distribution, thus the lower the variance of the weights,
the more efficiently the discrete distribution is in describing the continuous distribution. 
A common measure of "Particle Degeneracy" is the effective number of particles, described
in (Bergman "Navigation and Tracking Applications", 1999, J S Liu and R Chen "Sequential 
Monte Carlo Methods for Dynamical Systems", 1998), which is based on the "true weight"
of each particle. Of course the true weight is unknown, so a heuristic approximating 
$N_{eff}$ is used:
\begin{equation}
\label{neff}
\hat{N}_{eff} \approx \frac{N_p}{\sum_{i=1}^{N_p} w_i^2}
\end{equation}
Any quick run of a particle filter will reveal that unless the prior is particularly accurate,
$N_{eff}$ drops precipitously.  To alleviate this problem
a common technique known as resampling must be applied. The idea of re-sampling is to 
draw from the approximate posterior, thus generating a replica of the posterior with 
a support more suited to the distribution. Thus, if weights are all set to $1/N_p$, and 
$N_p$ points are drawn from the posterior,
\begin{equation}
\hat{\chi}_j \sim \left(\sum_{i=1}^{N_p} w_i(t)\delta(x - x_i(t))\right), j \in \{1, ..., N_p\}
\end{equation}
then $\hat{\chi} \sim \hat{x}$ should hold. Unfortunately, this isn't necessarily the truth: since the support is
still limited to the original particles, the number of unique particles can only go down.
This effect, often dubbed "particle impoverishment" can result in excessive quantization
errors in the final distribution. However, there is a solution. Instead of sampling from the
discrete distribution, a smoothing kernel is applied, and $\hat{\chi}_j$ are drawn from
that distribution. Because the distribution is continuous, there is no way for a collapse
of the particles to occur. The question then, is how to decide on the smoothing kernel. 
Often times the easiest way to sample from the continous distribution is to break the 
re-sampling down into two steps. First a member of the discrete distribution is randomly
selected based on the weights, and then based on the smoothing a nearby state variable 
is selected. The process of the selection will be defined as:
\begin{equation}
\chi_i = x_i + h\sigma \epsilon
\end{equation}
Where $h$ is the bandwidth, $\sigma$ is the standard deviation such that $\sigma \sigma^T = cov(x)$
and $\epsilon$ is drawn from the chosen kernel.
It has been proven that when all the elements of the mixture
have the same weight, as is the case after basic resampling, the kernel that minimizes the 
MSE between the estimated and true posterior is the Epanechnikov Kernel (cite Improving Regularised
Particle Filters, C Musso, N Oudjane and F LeGrand). 
\begin{equation}
K = \left\{
\begin{array}{lr}
\frac{n_x+2}{2c_{n_x}}(1-\|x\|^2) & if\ \|x\| < 1\\
0 & otherwise
\end{array}\right.
\end{equation}
%<more here>

@@

Because $\nu_d$ is a Weiner process, then $\nu_d(t) - \nu_d(t-1)$ is simply a Gaussian step. If
$\nu_d$ is some other additive process, the difference will still be one of a few stable
distributions. If $\nu_y$ is i.i.d. then the resulting distribution will still be zero mean
with a maximum variance of twice the original variance. All the assumptions made originally
for the particle filter hold, and all of the parameters may be distringuished based on
the step sizes, thus it is not unreasonable to attempt to match the string of step sizes
rather than string of direct readings.

<frequncy response graphs, highlighting noise frequency range and signal frequency range>

\subsubsection{Weighting Function}
Because $\hat{Pr}(y(T) | x(T))$, what I will call the weighting funciton,
is based on an unknown distribution, it is necessary to decide on a function
that will approximate $\hat{Pr}(y(t) | x(T))$. Obviously the function, $\omega(y(t), f(x(t))$
needs to centered at zero and have a scale comparable to the signal levels.
While a Gaussian function is the natural choice, we also wanted to try a distribution
with wider tails, so that outliers don't completely destroy particle's weights.
Therefore, we tried three weighting functions; Gaussian, exponential and 
the Cauchy distribution. The standard deviations (or scale) was set to 
$\sigma_y/5$, where $\sigma_y$ is the variance of all $y[0:\infty]$. Of course 
since the particle filter requires a weighting function to run, this means that
before the particle filter starts all the measurements have to be in. In cases where
this is impossible, a heuristic based on a small sample may work just as well.

The shape of the weighting function is extremely important, because it essentially decides
the rejection rate of particles. A very thin gaussian probability 
distribution function has nice properties, but thin tails. As a result, large
outliers in the measurement vector could easily force all the particles to
have near 0 weights, thus forcing the particle filter to converge improperly. 
On the other end of the spectrum, a cauchy PDF, has relatively fat tails, and
may not weight central particles high enough, preventing the particles from 
converging at a reasonable rate. Exponetials have the benefit of having an 
extremely smooth drop to zero, and a slope of 1 at the origin. Having a non-zero
slope at the origin is beneficial because it discriminates all the way up until
the measured and predicted $y$ are the same. The importance of the weighting function
cannot be overstated, as this is the primary factor in deciding the rate at
which the particle filter converges.

Obviously the optimal $\hat{Pr}(y(T) | x(T))$ is the true $Pr(y(T) | x(T))$; however
since that is unknown, I tested multiple different distributions. As stated previous,
because the exact scale of $Pr(y(T) | x(T))$ it not even known, the exponential distribution
has the added benefit of having a negative slope all the way form zeo to infinity. Thus
even if the scale is completely wrong, particles will still be well differentiated.

<Q-Q plot of real fmri data with Gaussian, DC>
<Q-Q plot of real fmri data with Gaussian, deltas>

@@

\begin{equation}
v_y(t)  = f(v_y(t-1, u(t)) =  \begin{cases} 
        u(t)G & \text{ if }  u(t)G-v_y \ge v_t\\
        v_y(t-1)\left(1 - \frac{1}{R_mC}\right) & \text{ if }  u(t)G-v_y < v_t
    \end{cases} 
\end{equation}

@@

where $\Gamma$ is the gamma function.
The in some ways the margin for error is decided by the weighting function, which
here will define as $W(V_y, v_y)$, where $V_y$ is the actual measurement, and 
$v_y$ is the estimate by a particular particle. The choice of this function is difficult,
and although the Gaussian is typically used, we found the exponential helpful
in dealing with particle deprivation.  The algorithm will then look like the following,

\begin{algorithmic}
\STATE Initialize $N_p$ Particles
\FOR{$i$ in $N_p$}
    \STATE $G \sim Gamma(\frac{\mu^2_G}{\sigma^2_G}, \frac{\sigma^2_G}{\mu_G})$
    \STATE $v_t \sim Gamma(\frac{\mu^2_{v_t}}{\sigma^2_{v_t}}, \frac{\sigma^2_{v_t}}{\mu_{v_t}})$
    \STATE $C \sim Gamma(\frac{\mu^2_C}{\sigma^2_C}, \frac{\sigma^2_C}{\mu_C})$
    \STATE $R_m \sim Gamma(\frac{\mu^2_R}{\sigma^2_R}, \frac{\sigma^2_R}{\mu_R})$
    \STATE $v_y = 0$, (Assume the system has been off for a long time)
    \STATE let $X_i(0) = \{G, v_t, C, R_m, v_y\}$
    \STATE let $w_i(0) = 1$ or to make a flat prior, $w_i(0) = \frac{1}{Pr(X_i(0))}$ 
\ENDFOR
\FOR{$t$ in Set of Measurement Times}
    \FOR{$i$ in $N_p$}
        \STATE $v_y(t) = f(v_y(t-1, u(t))$
        \STATE $w_i(t) = w_i(t-1)W(V_y(t), v_y(t))$
    \ENDFOR
\ENDFOR

>>>>

In order to characterize the noise in the FMRI we took a few 
sample time-series during resting state, which should theoretically be all noise and 
used Q-Q plots to analyze the type of noise present. \ref{fig:QQDC} shows the 
results of a Q-Q plots between the samples and a Gaussian of mean and variance based
on the samples. Because it the noise is often considered to have Weiner noise with
Gaussian steps, \ref{fig:QQDelta} shows a Q-Q plot using the steps, rather than 
absolute values.
\begin{figure}
\label{fig:QQDC}
\caption{Q-Q Plots of resting state data}
\end{figure}

\begin{figure}
\label{fig:QQDelta}
\caption{Q-Q Plots of resting state data, using the steps as the samples}
\end{figure}

Finally, to determine whether much of the autocorrelation was removed by taking
steps we plotted the autocorrelation of each sample location, before for the
actual value (\ref{fig:CorrDC}), and the steps between values (\ref{fig:CorrDelta}).
Ideally, if the steps are Gaussian, they should be independent and have a
flat autocorrelation; although in real life this is far too much to expect.

\begin{figure}
\label{fig:CorrDC}
\caption{Autocorrelation of the absolute value of various time-series}
\end{figure}

\begin{figure}
\label{fig:CorrDelta}
\caption{Autocorrelation of the steps of various time-series}
\end{figure}

<conclusion about the noise>

\subsection{Analysis of the BOLD Signal}
Several studies have endeavored to analyze the BOLD equations, which
we will review here. The most obvious analysis is to calculate the
steady state signal, when $u$ is held constant long enough for
all the transients to settle out.

\begin{eqnarray}
s_{ss} &=& 0 \nonumber \\
f_{ss} &=& \tau_f\epsilon u + 1\nonumber \\
v_{ss} &=& (\tau_f\epsilon u + 1)^\alpha\nonumber \\
q_{ss} &=& \frac{(\tau_f\epsilon u + 1)^\alpha}{E_0}(1-(1-E_0)^{1/(\tau_f\epsilon u + 1)})\nonumber \\
y_{ss} &=& V_0((k_1+k_2)(1-q_{ss}) - (k_2+k_3)(1-v_{ss}))
\label{steadystate}
\end{eqnarray}

[Image with two different $\alpha$'s]

[image comparing the results of 10\% changes in various signals]

@@

\subsection{General Linear Model}
The most used form of FMRI analysis is Statistical Parametric
Mapping, but is able to account for several different levels or types
of stimulus (see \cite{Hofmann1997}). By using hierarchical models
the output signal timeseries is considered the weighted
sum of the various input timeseries. Essentially every experimental factor
is considered as another level inputs (ex. different type of stimulus,
different run with the same patient, different patient). 
The equation for a general linear model is then
\begin{equation}
Y(t) = X(t)\beta + \epsilon(t)
\end{equation}
where $Y(t)$ is the smoothed or detrended timeseries of measurements,
$X(t)$ is a row vector of input, $\beta$ is a column vector of weights,
and $\epsilon$ is the error. Thus for every time, the measurement is
assumed to be a weighted sum of the columns of $X$ plus some error. The calculation
of $\beta$ is then performed using a maximum likelihood or gradient descent search 
to minimize the error.

@@

It is well known of course that a square wave stimulus does not result in a square wave
in the activation of brain regions. Thus, various methods are used to 
smooth the columns of $X$, and thus bandlimit the input. 
The best technique is convolving the stimulus input with a Hemodynamic 
Response Function (HRF), which mimics the basic shape of BOLD activation, including a delay
due to rise time and fall time. The downside of this method, however is that 
the shape of the Hemodynamic Signal is static, meaning the same Hemodynamic Function is
used for every region of the brain, and $Y(t)$ must linear combination 
of scalar the columns of $X$ to be appropriately identified. Additionally, 
it is well known that different Hemodynamic Response Functions are necessary for different 
regions of the brain. The "Canonical" HRF that is most used, has been fitted
for the visual cortex. Thus the shape of the HRF may differ significantly in
terms of onset and fall time in other areas of the brain. The inability to
fit parameters other than scale certainly hinders SPM's
ability to locate voxels that do not conform to the "canonical" hemodynamic 
response function. If there are different HRF's for different regions 
of the brain, might there also be more subtle differences even within
a particular region? It seems reasonable to think so. As a result, the 
most common use of SPM will be heavily biased toward the visual cortex.

@@

The GLM is extremely powerful at determining the linear dependence of
a set of regressors on the output. Unfortunately, there is significant evidence
that many of these dependencies are nonlinear, which means they may be
difficult or impossible to detect using linear techniques. This presents
a significant and unknown problem that is often left un-addressed in 
neural studies.

A static linear model is also unable to incorporate other forms of physiological
data that may be available. Combined FMRI
CBF or CBV imaging methods are not uncommon and they could shed much light on
neural activation. However, there is no way of actively combining that data 
into a static HRF. The fact of the matter is that the relation between the 
BOLD signal and CBV/CBF simply cannot be described by any linear relationship.
While it may be possible to determine some regionally varying set of parameters
for the HRF based on CBV/CBF, the lack of physiologically inspired parameters
impairs this. The importance of regional differences in the HRF cannot be
over-stated; it is well known that capillary beds are far from uniformly distributed
and thus blood perfusion and regional oxygenation cannot possibly be uniform
either.

Because of these limitations, its entirely possible that activation exists
in regions that SPM doesn't detect, but because the activation does not conform
to the set HRF, it is impossible for these signals to be detected. Perhaps
because of this, it is not uncommon for data to be thrown out or averaged
together in FMRI studies because 
no significant activation could be seen in a single run \cite{Riera2004}
\cite{Johnston2008}. This practice highlights 
limitations in strictly linear approaches; and suggests that a single HRF is 
insufficiently flexible to account for relatively common variations of neural activation.

@@

\begin{eqnarray}
X(t) &=& g(u(t), X(t-1))\\
Y(t) &=& h(X(t)
\end{eqnarray}

In order to estimate the posterior at $t$, a deterministic set of "sigma" points 
(often 2 per dimension, plus 1 at the mode of the multivariate distribution)
weighted according to a Gaussian estimate of $X(t-1)$ are passed through
the update equation. This set of points are then used to estimate the 
mean and covariance of $X(t)$. The benefit of this, is that it requires
no Jacobian and only a few extra calculations to get a decent esimate of
a posterior Gaussian. In the BOLD case, the set of equations we are modeling have no 
closed form solution, and finding the Jacobian is impossible without approximations. Although 
\cite{Riera2004}, \cite{Hu2009} mention a Jacobian of the BOLD response, this is
not strictly the case and is rather $\frac{\partial J}{\partial t}$
rather than a true Jacobian.  This is important because the Extended
Kalman filter depends on the Jacobian to map a Gaussian through the 
advancement of time. Thus the popular Extended Kalman Filter won't work
in this case, whereas the Unscented Kalman Filter still does. In fact
\cite{Hu2009} uses the UKF to perform a similar type of analysis to
the one performed in this work.

@@

\subsection{Conclusion}

@@

\subsubsection{Resampling}
\label{Sub:Resampling}
As a consequence 
of the wide prior distribution (required for a proper discretization of a continuous
distribution), there will be many particles with insignificant weights. While this does help
describe the tails of the distribution very well, it means that only a small portion of the
computation will be spent describing the most probable region. Ideally every particle would 
equally decrease the entropy of the distribution, thus the lower the variance of the weights,
the more efficiently the discrete distribution is in describing the continuous distribution. 
A common measure of "Particle Degeneracy" is the effective number of particles, described
in (Bergman "Navigation and Tracking Applications", 1999, J S Liu and R Chen "Sequential 
Monte Carlo Methods for Dynamical Systems", 1998), which is based on the "true weight"
of each particle. Of course the true weight is unknown, so a heuristic approximating 
$N_{eff}$ is used:
\begin{equation}
\label{neff}
\hat{N}_{eff} \approx \frac{N_p}{\sum_{i=1}^{N_p} w_i^2}
\end{equation}
Any quick run of a particle filter will reveal that unless the prior is particularly accurate,
$N_{eff}$ drops precipitously.  To alleviate this problem
a common technique known as resampling must be applied. The idea of re-sampling is to 
draw from the approximate posterior, thus generating a replica of the posterior with 
a support more suited to the distribution. Thus, if weights are all set to $1/N_p$, and 
$N_p$ points are drawn from the posterior,
\begin{equation}
\hat{\chi}_j \sim \left(\sum_{i=1}^{N_p} w_i(t)\delta(x - x_i(t))\right), j \in \{1, ..., N_p\}
\end{equation}
then $\hat{\chi} \sim \hat{x}$ should hold. Unfortunately, this isn't necessarily the truth: since the support is
still limited to the original particles, the number of unique particles can only go down.
This effect, often dubbed "particle impoverishment" can result in excessive quantization
errors in the final distribution. However, there is a solution. Instead of sampling from the
discrete distribution, a smoothing kernel is applied, and $\hat{\chi}_j$ are drawn from
that distribution. Because the distribution is continuous, there is no way for a collapse
of the particles to occur. The question then, is how to decide on the smoothing kernel. 
Often times the easiest way to sample from the continous distribution is to break the 
re-sampling down into two steps. First a member of the discrete distribution is randomly
selected based on the weights, and then based on the smoothing a nearby state variable 
is selected. The process of the selection will be defined as:
\begin{equation}
\chi_i = x_i + h\sigma \epsilon
\end{equation}
Where $h$ is the bandwidth, $\sigma$ is the standard deviation such that $\sigma \sigma^T = cov(x)$
and $\epsilon$ is drawn from the chosen kernel.
It has been proven that when all the elements of the mixture
have the same weight, as is the case after basic resampling, the kernel that minimizes the 
MSE between the estimated and true posterior is the Epanechnikov Kernel (cite Improving Regularised
Particle Filters, C Musso, N Oudjane and F LeGrand). 
\begin{equation}
K = \left\{
\begin{array}{lr}
\frac{n_x+2}{2c_{n_x}}(1-\|x\|^2) & if\ \|x\| < 1\\
0 & otherwise
\end{array}\right.
\end{equation}
%<more here>

@@

Because $\nu_d$ is a Weiner process, then $\nu_d(t) - \nu_d(t-1)$ is simply a Gaussian step. If
$\nu_d$ is some other additive process, the difference will still be one of a few stable
distributions. If $\nu_y$ is i.i.d. then the resulting distribution will still be zero mean
with a maximum variance of twice the original variance. All the assumptions made originally
for the particle filter hold, and all of the parameters may be distinguished based on
the step sizes, thus it is not unreasonable to attempt to match the string of step sizes
rather than string of direct readings.

<frequency response graphs, highlighting noise frequency range and signal frequency range>

\subsubsection{Weighting Function}
Because $\hat{Pr}(y(T) | x(T))$, what I will call the weighting function,
is based on an unknown distribution, it is necessary to decide on a function
that will approximate $\hat{Pr}(y(t) | x(T))$. Obviously the function, $\omega(y(t), f(x(t))$
needs to be centered at zero and have a scale comparable to the signal levels.
Obviously if the actual noise present in $y(t)$ were to be known, then that would
be the best distribution for $\Omega$. In that case, particles that fell
far out on that distribution would be statistically impossible representations
of the system, and it would be completely reasonable to throw such particles away.
While a Gaussian function is the natural choice, because this distribution
and weight are unknown we wanted to try distributions
with wider tails, so that outliers don't completely destroy particle's weights
(and thus convergence proceeds more slowly).

Another natural choice might be one of the robust estimator weight functions, for
example the Huber or bi-square. For the purpose of this work we will stick
with long tailed distributions, however it is worth noting that long tails
may not be the optimal choice in all, or even this situation. The justification
for long tailed distributions is that we believe the noise to be long tailed,
and the variance of the noise is not well known.

Therefore, we tried three weighting functions based on three distributions: Gaussian, 
Laplace and the Cauchy. The standard deviation of the distribution is extremely
important to the convergence of particle filter. A standard deviation that is 
too large will not allow the distribution to converge in any reasonable number of 
measurements. A standard deviation below the standard deviation of the noise 
will cause the algorithm to throw out perfectly acceptable particles. 
The weighting function ultimately will shape the output distribution, $P[y]$, into that
distribution, however the distribution of $x$ will still approach a reasonable
estimate of its true distribution. Even an overly wide weighting function, will
allow the Gaussian Mixture estimate of $X$ to converge to the correct location 
parameters of the "real" posterior distribution, though the scale parameters may 
be overly large.

A reasonable method of setting the standard deviation of $\Omega$ may be by 
taking a small sample from "resting" data and using the sample standard deviation.
Since this is the first attempt at using particle filters for modeling the 
BOLD model, in this work we set the standard deviation manually at <weight standard dev>,
because it gives a more consistency and control. Of course this could be taken
further, by testing the sample data against a set of stock distributions
and choosing the best fit. This of course depends on having enough samples
to make a reasonable inference, which may not always exist.

@@

\begin{equation}
v_y(t)  = f(v_y(t-1, u(t)) =  \begin{cases} 
        u(t)G & \text{ if }  u(t)G-v_y \ge v_t\\
        v_y(t-1)\left(1 - \frac{\delta t}{R_mC}\right) & \text{ if }  u(t)G-v_y < v_t
    \end{cases} 
\end{equation}

@@

where $\Gamma$ is the gamma function.
The in some ways the margin for error is decided by the weighting function, which
here will define as $W(V_y, v_{yi})$, where $V_y$ is the actual measurement, $v_y$ is the 
estimate based on all the particles, and 
$v_{yi}$ is the estimate by a particular (i$^{\text{th}}$) particle. The choice of this function is difficult,
and although the Gaussian is typically used, we found the exponential helpful
in dealing with particle deprivation.  The algorithm will then look like the following,

\begin{algorithmic}
\STATE Initialize $N_p$ Particles:
\FOR{$i$ in $N_p$}
    \STATE $G \sim Gamma(\frac{\mu^2_G}{\sigma^2_G}, \frac{\sigma^2_G}{\mu_G})$
    \STATE $v_t \sim Gamma(\frac{\mu^2_{v_t}}{\sigma^2_{v_t}}, \frac{\sigma^2_{v_t}}{\mu_{v_t}})$
    \STATE $C \sim Gamma(\frac{\mu^2_C}{\sigma^2_C}, \frac{\sigma^2_C}{\mu_C})$
    \STATE $R_m \sim Gamma(\frac{\mu^2_R}{\sigma^2_R}, \frac{\sigma^2_R}{\mu_R})$
    \STATE $v_y = 0$, (Assume the system has been off for a long time)
    \STATE let $X_i(0) = \{G, v_t, C, R_m, v_y\}$
    \STATE let $w_i(0) = 1$ or to make a flat prior, $w_i(0) = \frac{1}{Pr(X_i(0))}$ 
\ENDFOR
\STATE Run the Filter:
\FOR{$t$ in Set of Measurement Times}
    \FOR{$i$ in $N_p$}
        \STATE $v_{yi}(t) = f(v_{yi}(t-1), u(t))$
        \STATE (All other members of $X_i(t)$ remain the same)
        \STATE $w_i(t) = w_i(t-1)W(V_y(t), v_y(t))$
    \ENDFOR
\ENDFOR
\end{algorithmic}

Initially the particles will have the same output, $0$, however, as $u(t)$
changes, the response of each particle to that input will result in different
outputs. Particles that have a $v_{yi}$ near $V_y$ will be weighted higher,
and others farther away will be weighted lower. As the particle filter
runs, weights will compound resulting in a distribution that asymptotically
approaches the true joint distribution of the $X(t)$.  Of course, as we
mentioned in \ref{Sub:Resampling}, particles weighted zero do not significantly
contribute to the empirical distribution, so re-sampling may be necessary.
